---
title: "Exploratory Data Analysis of ISPARK Dataset"
output:
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ggplot2)
library(lubridate)
parkinformations <- readRDS("ispark-parkbilgileri.RDS") %>% select(-c(Distance,BosKapasite))
parkcapacitylog <- readRDS("parkcapacitylog.RDS")
```

## Dataset

We have two tables in our dataset. First table contains information about ISPARK park areas. Second table contains hourly percentage of occupancy of these park areas between 19 November 2020 and 2 December 2020. We have collected this datas by using IBB Open Data API. You can also access this API from [here.](https://api.ibb.gov.tr/ispark/Park)

### Park Informations Table

Park Information table contains 9 variables and 577 observations. Descriptions of columns is below.

* `ParkID`: Unique identifier of the parking place
* `ParkAdi`: Name of the parking place
* `Latitude`: Latitude of the parking place
* `Longitude`: Longitude of the parking place
* `Kapasitesi`: Full capacity of the parking space when it is empty
* `ParkTipi`: Types of parking ways of the parking space; Yolüstü – Acik-- Kapali
* `Ilce`: District which parking space in
* `UcretsizParklanmaDk`:Free parking duration in minutes
* `AylıkAbonelikUcreti`:Monthly subscription fee (₺)

#### Summary of Dataset

```{r parkinfosum}
summary(parkinformations)
```


### Hourly Percentage of Occupancy Table

Hourly Percentage of Occupancy table contains 4 variables and `r nrow(parkcapacitylog)` observations. Descriptions of columns is below.

* `ParkID`: Unique identifier of the parking place
* `ParkAdi`: Name of the parking place
* `DolulukYuzdesi`: Percentage of occupancy
* `OlcumZamanı`: Observation date and time

#### Summary of Dataset

```{r parkcapsum}
summary(parkcapacitylog)
```

### Missing Values

We don't have any missing values in our tables as you can see below. We think this is because we use API data that is cleaned and submitted by others.

Missing values in Park Information table by variable

```{r}
sapply(parkinformations, function(x) sum(is.na(x)))
```

Missing values in Hourly Percentage of Occupancy table by variable

```{r}
sapply(parkcapacitylog, function(x) sum(is.na(x)))
```

## Statistical Examination

### Park Type Summary Table

According to their types, we can see the count, total capacity and avg. monthly price in below table.

```{r message=FALSE, warning=FALSE}
parkinfo_tip <- parkinformations %>% group_by(ParkTipi) %>% summarize("Toplam Kapasite"=sum(Kapasitesi), "Ortalama Kapatise"=mean(Kapasitesi), "Ortalama Ücretsiz Park Süresi" = mean(UcretsizParklanmaDk), "Ortalama Aylık Ücret" = mean(AylıkAbonelikÜcreti))
parktipi_adet <- parkinformations %>% group_by(ParkTipi) %>% count(ParkTipi, sort = TRUE)
#parkinfo_tip %>% left_join(parktipi_adet) 
parktipi_adet %>% left_join(parkinfo_tip) %>% rename("Park Adeti"=n) 
```

### Park Count by District

Top 10 district who has most park count are as follows;

```{r message=FALSE, warning=FALSE}
parkyer_adet <- parkinformations %>% group_by(Ilce) %>% count(Ilce, sort = TRUE) %>% top_n(10, n)
parkyer_adet
```

```{r}
parktipi_adet <- parkinformations %>% group_by(ParkTipi) %>% count(ParkTipi, sort = TRUE)

```

### Boxplot of Avg. monthly price for each type of `Park Type`

We took out  2 of the park types because their quantity was lower than the others so they were misleading.
In conclusion,
* 'Open Park' has a lower mean of monthly price than 'Close Park' and 'Road Park'. 
* Close Park's monthly price has outliers. But despite those outliers, its monthly prices varies the least.

```{r}
parkinformations %>% filter(ParkTipi %in% c("AÇIK OTOPARK","KAPALI OTOPARK","YOL ÜSTÜ")) %>% ggplot(., aes(ParkTipi,AylıkAbonelikÜcreti)) +
  geom_boxplot()
```

### Hourly Occupancy Rates

```{r message=FALSE, warning=FALSE}
park_capacity_hourly <- parkcapacitylog %>%
  group_by(hour = lubridate::hour(OlcumZamanı)) %>%
  summarize( avd_doluluk_orani =mean(DolulukYuzdesi))
ggplot(park_capacity_hourly, aes(hour, avd_doluluk_orani)) +
  geom_line() +
  labs(x="Hour", y="Average Occupancy Rate") +
  labs(title = "Average Hourly Parking Occupancy Rate in Istanbul")
```


In above line graph, average parking occupancy rate during daytime is less than nights. Parking occupancy rate is starting to decrease when working hours started and continue to decrease until approximately 15:00 pm. and it sharply increases by closing hours of working day. The graph shows that 'İspark' stations in Istanbul mostly used during nights.


```{r message=FALSE, warning=FALSE}
mergedtable <- merge(x = parkcapacitylog, y = parkinformations, by = "ParkID", all.x = TRUE)
merged_park_capacity_hourly <- mergedtable %>%
  group_by(hour = lubridate::hour(OlcumZamanı), ParkTipi) %>%
  summarize( avd_doluluk_orani =mean(DolulukYuzdesi)) %>% filter(ParkTipi != "NA")
ggplot(merged_park_capacity_hourly, aes(hour, avd_doluluk_orani, color=ParkTipi)) +
  geom_line() +
  labs(x="Hour", y="Average Occupancy Rate") +
  labs(title = "Hourly Parking Occupancy Rate by Park Type in Istanbul")
```


In all park types, the average occupancy rate decreases during daytime In Istanbul. In average, it is more probably that chance to find empty parking space in parking garage than open parking lots and parking lots near road.  


### Price Histogram

The histogram below shows us the count of monthly membership prices for different parking types. It has been seen that monthly price of 400 TL and 100 TL are the least common prices. Count of open parks priced 200 TL and   count of parks with 0 price are almost identical to each other. According to the histogram We can say that the count of parks it mostly scattered around 200 TL. Free parking lots which doesn't allow monthly parking are the second most common park type. Also the measured data is a bi-modal distribution with two peaks.


```{r}
ggplot(parkinformations, aes(x=AylıkAbonelikÜcreti, fill=ParkTipi))+
  geom_histogram(bins = 6, binwidth = 100)+
  labs(title="Price Histogram",x ="Price", y = "Count")
```

