---
title: "Final Report of ISPARK Dataset"
output:
  html_document: 
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ggplot2)
library(lubridate)
parkinformations <- readRDS("ispark-parkbilgileri.RDS") %>% select(-c(Distance,BosKapasite))
parkcapacitylog <- readRDS("parkcapacitylog.RDS")
```

# Key Takeaways

* Total number of minibus and Taxi type of parks in Istanbul make the %2 of the total parks.
* Fatih (%17) - Besiktas (%11) - Sisli (%11) - Bakirköy (%6.58) - Kadiköy (%6.58) are the top 5 districts that have the most park number %52 of the total number. But on the other hand capacity-wise Eyup has the biggest park capacity. We can say touristic areas has an advantage in terms of parking.
* %33.96 of the parks are free. And On-Road or open parks are more likely to be free.
* At 15:00 occupancy rate is the lowest for all park types expect Taxi parks. And closed parks are most likely to have empty space than others.

# Our Mission

Understanding the relation between the park capacity and the other metric. Creating shiny app that recommends best suitable parking spots with given limits. You can check our app from [here.](https://alihan.shinyapps.io/ISPARK-Shiny/)

# Brief Information About Dataset

We have two tables in our dataset. First table contains informations about ISPARK park areas. Second table contains hourly percentage of occupancies of these park areas between 19 November 2020 and 2 December 2020. We have collected this data by using IBB OpenData API. You can also access this API from [here.](https://api.ibb.gov.tr/ispark/Park)

# Important Tables & Plots

## Park Informations Table

Park Informations table contains 9 variables and 577 observations. Descriptions of columns is below.

* `ParkID`: Unique identifier of the parking place
* `ParkAdi`: Name of the parking place
* `Latitude`: Latitude of the parking place
* `Longitude`: Longitude of the parking place
* `Kapasitesi`: Full capacity of the parking space when it is empty
* `ParkTipi`: Types of parking ways of the parking space; Yolüstü – Acik-- Kapali
* `Ilce`: District which parking space in
* `UcretsizParklanmaDk`:Free parking duration in minutes
* `AylıkAbonelikUcreti`:Monthly subscription fee (₺)

```{r parkinfosum}
summary(parkinformations)
```



## Hourly Percentage of Occupancy Table

Hourly Percentage of Occupancy table contains 4 variables and `r nrow(parkcapacitylog)` observations. Descriptions of columns is below.

* `ParkID`: Unique identifier of the parking place
* `ParkAdi`: Name of the parking place
* `DolulukYuzdesi`: Percentage of occupancy
* `OlcumZamanı`: Observation date and time


```{r parkcapsum}
summary(parkcapacitylog)
```


## Park Type Summary Table

According to their types, we can see the count, total capacity and avg. monthly price in below table.

```{r message=FALSE, warning=FALSE}
parkinfo_tip <- parkinformations %>% group_by(ParkTipi) %>% summarize("Toplam Kapasite"=sum(Kapasitesi), "Ortalama Kapatise"=mean(Kapasitesi), "Ortalama Ücretsiz Park Süresi" = mean(UcretsizParklanmaDk), "Ortalama Aylık Ücret" = mean(AylıkAbonelikÜcreti))
parktipi_adet <- parkinformations %>% group_by(ParkTipi) %>% count(ParkTipi, sort = TRUE)
#parkinfo_tip %>% left_join(parktipi_adet) 
parktipi_adet %>% left_join(parkinfo_tip) %>% rename("Park Adeti"=n) 
```


```{r}
parktipi_adet <- parkinformations %>% group_by(ParkTipi) %>% count(ParkTipi, sort = TRUE)

```

## Boxplot of Avg. monthly price for each type of `Park Type`

We took out  2 of the park types because their quantity was lower than the others so they were misleading.
In conclusion,
* 'Open Park' has a lower mean of monthly price than `Close Park` and `Road Park`. 
* Close Park's monthly price has outliers. But despite those outliers, its monthly prices varies the least.

```{r, figures-side, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
parkinformations %>% filter(ParkTipi %in% c("AÇIK OTOPARK","KAPALI OTOPARK","YOL ÜSTÜ")) %>% ggplot(., aes(ParkTipi,AylıkAbonelikÜcreti)) +
  geom_boxplot() +
  labs(title =" Avg. Monthly Price - All ", x = "Park Tipi", y = "Avg. Price") 

parkinformations %>% filter(ParkTipi %in% c("AÇIK OTOPARK","KAPALI OTOPARK","YOL ÜSTÜ")) %>% filter(AylıkAbonelikÜcreti != 0) %>%
  ggplot(., aes(ParkTipi,AylıkAbonelikÜcreti)) +
  geom_boxplot() +
  labs(title =" Avg. Monthly Price - 0 Excluded ", x = "Park Tipi", y = "Avg. Price") 
```

## Hourly& Daily Occupancy Rate

The line below shows us average hourly parking occupancy Rates by park types of ‘Ispark’ stations. In all park types, average parking occupancy rate during daytime (around 15:00) is less than nights. In other words, in all park types, park stations mostly used during night. 
Another conclusion we can draw is, in average, it is more probably that chance to find empty parking space in parking garage than open parking lots and parking lots near road.
We can say on week days occupancy is lower than weekend days.
Taxi Parks, seems %100 occupied all the time. Meaning IBB might want it to seem full all the time in dataset.


```{r hourday, figures-side, fig.show="hold", out.width="50%", message=FALSE, warning=FALSE}
mergedtable <- merge(x = parkcapacitylog, y = parkinformations, by = "ParkID", all.x = TRUE)
merged_park_capacity_hourly <- mergedtable %>%
  group_by(hour = lubridate::hour(OlcumZamanı), ParkTipi) %>%
  summarize( avd_doluluk_orani =mean(DolulukYuzdesi)) %>% filter(ParkTipi != "NA")
ggplot(merged_park_capacity_hourly, aes(hour, avd_doluluk_orani, color=ParkTipi)) +
  geom_line() +
  labs(x="Hour", y="Average Occupancy Rate") +
  labs(title = "Hourly Parking Occupancy Rate by Ispark Types")


merged_park_capacity_daily <- mergedtable %>% 
  group_by(wday = lubridate::wday(OlcumZamanı), ParkTipi) %>%
  summarize( avd_doluluk_orani =mean(DolulukYuzdesi)) %>% filter(ParkTipi != "NA")
  ggplot(merged_park_capacity_daily, aes(wday, avd_doluluk_orani, color=ParkTipi)) +
  geom_line() +
  labs(x="Week Day", y="Average Occupancy Rate") +
  labs(title = "Daily Parking Occupancy Rate by Ispark Types")


```


## Price Plots

We have 196 free, 381 paid parks in Istanbul.

%52 of the free parks are On Road Type of park. %40.8 of the free parks are Open Parks.

```{r ucretdurumu, figures-side, fig.show="hold", out.width="50%", message=FALSE, warning=FALSE}
ucretdurumu <- parkinformations %>% transmute(ucretDurumu = ifelse(AylıkAbonelikÜcreti == 0,"Free","Not Free"))
ucretdurumutab <- ucretdurumu %>% group_by(ucretDurumu) %>% count(ucretDurumu)

ggplot(ucretdurumutab, aes(x=reorder(ucretDurumu, n, fill=ucretDurumu), y=n))+
  geom_bar(stat="identity", position = "dodge")+
  labs(title="Free Parks",x ="Free/ Not Free", y = "Count")

ucretdurumu2 <- parkinformations %>% transmute(ParkTipi, ucretDurumu = ifelse(AylıkAbonelikÜcreti == 0,"Free","Not Free"))
ucretdurumutab2 <- ucretdurumu2 %>% group_by(ParkTipi, ucretDurumu) %>% count(ucretDurumu)

ggplot(ucretdurumutab2, aes(x=reorder(ucretDurumu, n), y=n, fill=ParkTipi))+
  geom_bar(stat="identity", position = "dodge")+
  labs(title="Free/ Not Free Status by Park Type",x ="Free/ Not Free", y = "Count")

```


In below table, we can see that;

* For on-road parks most frequent monthly price is between 200 - 300 TL.

* For open parks most frequent monthly price is between 100 - 200 TL.

* For closed parks most frequent monthly price is close to 200 TL.


```{r, out.width="60%", message=FALSE, warning=FALSE}
parkinformations %>% filter(AylıkAbonelikÜcreti != 0) %>%
  ggplot(aes(x=AylıkAbonelikÜcreti, color=ParkTipi))+
  geom_freqpoly(binwidth = 50, position="dodge")+
  labs(title="Monthly Membership Prices Histogram",x ="Price", y = "Count")
```

## Total Park Capacities of Districts 

In this graph if we look at the total park capacities of districts. We can see Open Parks have more capacity than other types of parks. And Eyup has the biggest park capacity.

```{r message=FALSE, warning=FALSE}
IlceToplamlar <- parkinformations %>% group_by(Ilce) %>% summarise(IlceKapasite=sum(Kapasitesi))

parkinformations %>% group_by(Ilce,ParkTipi) %>% summarise(Kapasite=sum(Kapasitesi)) %>% 
  right_join(., IlceToplamlar, by="Ilce") %>% 
  ggplot(., aes(x=reorder(Ilce,-IlceKapasite), y=Kapasite, fill=ParkTipi)) +
       geom_bar(position = "stack", stat = "identity", width = 0.5) +
       theme(axis.text.x = element_text(angle=65, vjust=0.6)) +
       xlab("District") +
       ylab("Park Capacity")+
      facet_grid(ParkTipi ~ .)

```

## Frequency of the Parks

If we look at the frequencies of the parks Fatih district has the highest frequency. After that Beşiktaş and Şişli districts are almost equal and has the second highest frequency. Arnavutköy and Sultangazi are almost equal and has the lowest frequency.

```{r, figures-side, fig.show="hold", out.width="50%", message=FALSE, warning=FALSE}
parkinformations %>% group_by(Ilce) %>% count() %>% 
       ggplot(mapping = aes(x = reorder(Ilce, n), y=n)) + geom_bar(stat = "identity", fill="green") +
       labs(title="Frequency of Park Areas")  +
       coord_flip() + 
       ylab("Park Area Count") +
       xlab("District") +
       theme_minimal()
```



