---
title: "Final Report of ISPARK Dataset"
output:
  html_document: 
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ggplot2)
library(lubridate)
parkinformations <- readRDS("ispark-parkbilgileri.RDS") %>% select(-c(Distance,BosKapasite))
parkcapacitylog <- readRDS("parkcapacitylog.RDS")
```

# Key Takeaways

* Total number of minibus and Taxi type of parks in Istanbul make the %2 of the total parks.
* Fatih (%17) - Besiktas (%11) - Sisli (%11) - Bakirköy (%6.58) - Kadiköy (%6.58) are the top 5 districts that have the most park number %52 of the total number.
* Open park type has the lowest average monthly parking price among others.
* At 15:00 occupancy rate is the lowest for all park types expect Taxi parks.

# Our Mission

Understanding the relation between the park capacity and the other metric. Creating shiny app that recommends best suitable parking spots with given limits. You can check our app from [here.](https://alihan.shinyapps.io/ISPARK-Shiny/)

# Brief Information About Dataset

We have two tables in our dataset. First table contains informations about ISPARK park areas. Second table contains hourly percentage of occupancies of these park areas between 19 November 2020 and 2 December 2020. We have collected this data by using IBB OpenData API. You can also access this API from [here.](https://api.ibb.gov.tr/ispark/Park)

# Important Tables & Plots

## Park Informations Table

Park Informations table contains 9 variables and 577 observations. Descriptions of columns is below.

* `ParkID`: Unique identifier of the parking place
* `ParkAdi`: Name of the parking place
* `Latitude`: Latitude of the parking place
* `Longitude`: Longitude of the parking place
* `Kapasitesi`: Full capacity of the parking space when it is empty
* `ParkTipi`: Types of parking ways of the parking space; Yolüstü – Acik-- Kapali
* `Ilce`: District which parking space in
* `UcretsizParklanmaDk`:Free parking duration in minutes
* `AylıkAbonelikUcreti`:Monthly subscription fee (₺)

```{r parkinfosum}
summary(parkinformations)
```



## Hourly Percentage of Occupancy Table

Hourly Percentage of Occupancy table contains 4 variables and `r nrow(parkcapacitylog)` observations. Descriptions of columns is below.

* `ParkID`: Unique identifier of the parking place
* `ParkAdi`: Name of the parking place
* `DolulukYuzdesi`: Percentage of occupancy
* `OlcumZamanı`: Observation date and time


```{r parkcapsum}
summary(parkcapacitylog)
```


## Park Type Summary Table

According to their types, we can see the count, total capacity and avg. monthly price in below table.

```{r message=FALSE, warning=FALSE}
parkinfo_tip <- parkinformations %>% group_by(ParkTipi) %>% summarize("Toplam Kapasite"=sum(Kapasitesi), "Ortalama Kapatise"=mean(Kapasitesi), "Ortalama Ücretsiz Park Süresi" = mean(UcretsizParklanmaDk), "Ortalama Aylık Ücret" = mean(AylıkAbonelikÜcreti))
parktipi_adet <- parkinformations %>% group_by(ParkTipi) %>% count(ParkTipi, sort = TRUE)
#parkinfo_tip %>% left_join(parktipi_adet) 
parktipi_adet %>% left_join(parkinfo_tip) %>% rename("Park Adeti"=n) 
```


```{r}
parktipi_adet <- parkinformations %>% group_by(ParkTipi) %>% count(ParkTipi, sort = TRUE)

```

## Boxplot of Avg. monthly price for each type of `Park Type`

We took out  2 of the park types because their quantity was lower than the others so they were misleading.
In conclusion,
* 'Open Park' has a lower mean of monthly price than 'Close Park' and 'Road Park'. 
* Close Park's monthly price has outliers. But despite those outliers, its monthly prices varies the least.

```{r}
parkinformations %>% filter(ParkTipi %in% c("AÇIK OTOPARK","KAPALI OTOPARK","YOL ÜSTÜ")) %>% ggplot(., aes(ParkTipi,AylıkAbonelikÜcreti)) +
  geom_boxplot()
```

## Hourly Occupancy Rate

The line graph below shows us average hourly occupancy rates of ‘Ispark’ stations. It can be seen from the graph that average parking occupancy rate during daytime is less than nights. Parking occupancy rate is starting to decrease when working hours started and continue to decrease until approximately 15:00 pm. and it sharply increases by closing hours of working day. The graph shows that ‘Ispark’ stations mostly used during night.

```{r message=FALSE, warning=FALSE}
park_capacity_hourly <- parkcapacitylog %>%
  group_by(hour = lubridate::hour(OlcumZamanı)) %>%
  summarize( avd_doluluk_orani =mean(DolulukYuzdesi))
ggplot(park_capacity_hourly, aes(hour, avd_doluluk_orani)) +
  geom_line() +
  labs(x="Hour", y="Average Occupancy Rate") +
  labs(title = "Average Hourly Parking Occupancy Rate of Ispark")
```


The line below shows us average hourly parking occupancy Rates by park types of ‘Ispark’ stations. In all park types, the average occupancy rate decreases during daytime. In other words, in all park types, park stations mostly used during night. Another conclusion we can draw is, in average, it is more probably that chance to find empty parking space in parking garage than open parking lots and parking lots near road.


```{r message=FALSE, warning=FALSE}
mergedtable <- merge(x = parkcapacitylog, y = parkinformations, by = "ParkID", all.x = TRUE)
merged_park_capacity_hourly <- mergedtable %>%
  group_by(hour = lubridate::hour(OlcumZamanı), ParkTipi) %>%
  summarize( avd_doluluk_orani =mean(DolulukYuzdesi)) %>% filter(ParkTipi != "NA")
ggplot(merged_park_capacity_hourly, aes(hour, avd_doluluk_orani, color=ParkTipi)) +
  geom_line() +
  labs(x="Hour", y="Average Occupancy Rate") +
  labs(title = "Hourly Parking Occupancy Rate by Ispark Types")
```


## Price Histogram

The histogram below shows us the count of monthly membership prices for different parking types. It has been seen that monthly price of 400 TL and 100 TL are the least common prices. Count of open parks priced 200 TL and   count of parks with 0 price are almost identical to each other. According to the histogram We can say that the count of parks it mostly scattered around 200 TL. Free parking lots which doesn't allow monthly parking are the second most common park type. Also the measured data is a bi-modal distribution with two peaks.


```{r}
ggplot(parkinformations, aes(x=AylıkAbonelikÜcreti, fill=ParkTipi))+
  geom_histogram(bins = 6, binwidth = 100)+
  labs(title="Price Histogram",x ="Price", y = "Count")
```

## Total Park Capacities of Districts 

In this graph if we look at the total park capacities of districts.

```{r message=FALSE, warning=FALSE}
IlceToplamlar <- parkinformations %>% group_by(Ilce) %>% summarise(IlceKapasite=sum(Kapasitesi))

parkinformations %>% group_by(Ilce,ParkTipi) %>% summarise(Kapasite=sum(Kapasitesi)) %>% 
  right_join(., IlceToplamlar, by="Ilce") %>% 
  ggplot(., aes(x=reorder(Ilce,-IlceKapasite), y=Kapasite, fill=ParkTipi)) +
       geom_bar(position = "stack", stat = "identity", width = 0.5) +
       theme(axis.text.x = element_text(angle=65, vjust=0.6)) +
       xlab("Park Capacity") +
       ylab("District")

```

## Frequency of the Parks

If we look at the frequencies of the parks Fatih district has the highest frequency. After that Beşiktaş and Şişli districts are almost equal and has the second highest frequency. Arnavutköy and Sultangazi are almost equal and has the lowest frequency.

```{r message=FALSE, warning=FALSE}
parkinformations %>% group_by(Ilce) %>% count() %>% 
       ggplot(mapping = aes(x = reorder(Ilce, n), y=n)) + geom_bar(stat = "identity", fill="green") +
       labs(title="Frequency of Park Areas")  +
       coord_flip() + 
       ylab("Park Area Count") +
       xlab("District") +
       theme_minimal()
  
```



